---
title: "NY Times Article Search"
author: "Alice Friedman, DATA 607"
date: "October 26, 2018"
output: 
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
---

```{r, include=FALSE}
Sys.setenv(NYTIMES_KEY="XXXXXXXXXXX")
```

#Setup
```{r, message=FALSE, results='hide'}
library(httr)
library(jsonlite)
library(dplyr)
library(ggplot2)
```
This code will not run without first obtaining an API Key from NYT.

Once acquired, save your API key to R using ```Sys.setenv(NYTIMES_KEY="INSERT YOUR KEY HERE")```

Your key can be retrieved without revealing it through the R code by then using ```Sys.getenv("NYTIMES_KEY")```



#API structure

This API works by taking the base url "http://api.nytimes.com/svc/search/v2/articlesearch.json?q=" and searching based on certain parameters, which include:

- q (query term)

- begin_date (YYYYMMDD format for earliest date to search)

- end_date (YYYYMMDD format for latest date to search)

- fl (list of fields to return)

To search for specific fields for a specific query term or time, we must set parameters and then write a short piece of code which pastes those into a new URL.

##Setting Parameters
```{r}
# Need to use + to string together separate words
q_term <- "dragons" 
begin_date <- "201701011"
end_date <- "20180101"
```

##Assembling the url and making the initial query

First, we paste our parameters together with the API key to develop the URI which will return a JSON response.
```{r}
baseurl <- paste0(
  "http://api.nytimes.com/svc/search/v2/articlesearch.json?q=",
  q_term,
  "&begin_date=",
  begin_date,
  "&end_date=",
  end_date,
  "&facet_filter=true&api-key=",
  Sys.getenv("NYTIMES_KEY"))

initialQuery <- fromJSON(baseurl)
```

As we can see, this search returned 31 hits, but only the API returns information about only 10 articles at a time.

#Creating a function
To make this repeatable over different parameters, we can create a function to allow for a search of different search terms that will also:

- Convert results to a data frame

- Manipulate the search query to automatically download all pages of matching responses

- Merge the different responses into a single data frame. 

For relative simplicity, we will choose just three parameters: search terms, earliest search date, and latest search date, and will limit fields returned to just a few.

Note: This function will still require an API key to function. To set yours, enter your key in the following code:

```Sys.setenv(NYTIMES_KEY="INSERT-YOUR-KEY-HERE")```  

```{r, message=FALSE, results='hide'}
articleSearch <- function(search_terms, search_from_date, search_to_date) {
  baseurl <- paste0(
      "http://api.nytimes.com/svc/search/v2/articlesearch.json?q=",
      search_terms,
      "&begin_date=",
      search_from_date,
      "&end_date=",
      search_to_date)
  initialQuery <- fromJSON(paste0(
    baseurl, 
    "&facet_filter=true&api-key=",
    Sys.getenv("NYTIMES_KEY")))
  hits <- initialQuery$response$meta$hits
  numPages <- round((hits/10)-1) 
  pages <- list()
#This Sys.sleep is very important--otherwise you will get a 429 error
  Sys.sleep(1)
  for(i in 0:numPages){
      nytSearch <- fromJSON(
        paste0(
          baseurl, 
          "&page=", 
          i, 
          "&api-key=",
          Sys.getenv("NYTIMES_KEY")), 
        flatten = TRUE) %>% 
        data.frame() 
      message("Retrieving page ", i)
      message(Sys.time())
      pages[[i+1]] <- nytSearch 
      #A second Sys.sleep is needed in the for loop
      Sys.sleep(2)
  }
  allNYTSearch <- rbind_pages(pages)
}
```

##Running the Function with Different Parameters
Let's try running the function on a few different dessert trends and checking out some snippets.

```{r}
Sys.sleep(2)
cronuts2018 <- articleSearch("cronuts","20180101","20181027")
cronuts2018$response.docs.snippet[1:3]

Sys.sleep(2)
cupcakes2018 <- articleSearch("cupcakes","20180101","20181027")
cupcakes2018$response.docs.snippet[1:3]
```

#Visualizing the Data
To visualize the data, we will make a function that takes the search response data frame as an input.
```{r}
#Create a function
sumBarChart <- function(dataframe) {
  dataframe %>% 
  group_by(response.docs.type_of_material) %>%
  summarize(count=n()) %>%
  ggplot() +
    geom_bar(aes(y=count, 
               x=response.docs.type_of_material, 
               fill=response.docs.type_of_material), 
           stat = "identity") + 
    coord_flip() +
    labs(y="Number of Articles", x="Type of Coverage")+
    theme(legend.position="none")
}

sumBarChart(cronuts2018)+ggtitle("NY Times Cronut Coverage, 2018")
sumBarChart(cupcakes2018)+ggtitle("NY Times Cupcake Coverage, 2018")
```

Looks like cupcakes were a lot more popular in 2018 than cronuts!


#References

The above code is heavily indebted to the work of Jonathan D. Fitzgerald, published on January 25, 2018 on <a href="http://www.storybench.org/" target="_blank"><i>Storybench</i></a>, <a href="http://www.storybench.org/working-with-the-new-york-times-api-in-r/" target="_blank">"Working with The New York Times API in R"</a>